Random Forests, LAs

### 2024-06-18 src/la_RF_corrected.Rmd
corrected RF

@@

df9610 <- read_csv(data.path9610)

@@

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

# Complete panel
## REPLACE WITH 0.1 INSTEAD OF OF 0 FOR THE GROWTH
df <- df %>% filter(!is.na(ladcd)) %>% 
  complete(ladcd, year) %>% 
  replace(is.na(.), 0)

@@

# [reproducible results]
# Random Forest 
# 
# 5984 samples
#    6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 5055, 4690, 4732, 5070, 5040, 4718, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE        Rsquared   MAE       
#   2     variance    0.03227897  0.8023185  0.01991724
#   2     extratrees  0.03150004  0.8057960  0.01912348
#   4     variance    0.03271887  0.7985855  0.02025685
#   4     extratrees  0.03174791  0.8047090  0.01932473
#   6     variance    0.03358260  0.7879953  0.02080616
#   6     extratrees  0.03207532  0.8018495  0.01964516
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.

@@

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

@@

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

@@

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

@@

@@

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

@@

## Granger

```{r}
library(plm)
p.df <- pdata.frame(df, index= c("ladcd","year"))

test <- pgrangertest(n ~ n.l.slag, data = p.df, order = 1L) # *** reject the null that x does NOT cause y
# n ~ Lags(n, 1) + Lags(n.l.slag, 1), or
# n ~ Lags(n, 1) + Lags(n.slag, 2)
# n ~ Lags(n, 1)
test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)

test <- pgrangertest(n ~ n.slag, data = p.df, order = 1L)
# n ~ Lags(n, 1:1) + Lags(n.slag, 1:1)
# n ~ Lags(n, 1:1)
test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)

test <- pgrangertest(n.slag ~ n, data = p.df, order = 1L) # ** reject the null that x does NOT cause y
test$indgranger %>% as_tibble() %>% arrange(`p-value`)

p.df$n.London_dist = p.df$n.London/p.df$London
p.df$n.nearest.city_dist = p.df$n.nearest.city/p.df$dist

pgrangertest(n ~ n.London_dist, data = p.df, order = 1L)
pgrangertest(n ~ n.nearest.city_dist, data = p.df, order = 1L)

cor(p.df$n.London_dist, dplyr::lag(p.df$n.London_dist), use = "complete.obs")
```

>>>>

## Problem postcodes

@@

df9610 <- read_csv(data.path9610) %>% 
  filter(
      #V1.domain < m,             # for multiple pc
      V1.domain == n,             # for n == 1
      year > 1995)

# df9610 %>% filter(V1.domain == 1) %>% 
#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
#   filter(ladcd == "E08000006" & year == 2006) %>% 
#   group_by(pc) %>% 
#   summarise(n = n()) %>% 
#   arrange(desc(n))
# 
# df9610 %>% filter(pc == "M28 2SL" & year == 2006 & V1.domain == 1)

problem.pcs <- df9610 %>% filter(year > 1999 & year < 2011) %>% 
  group_by(pc, year) %>% 
  summarise(n = n()) %>%
  ungroup() %>% 
  # complete(pc, year) %>% 
  # replace(is.na(n), 0) %>% 
  arrange(desc(n)) %>% 
  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
  dplyr::select(pc, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010) %>% 
  slice_head(n=10) %>% 
  dplyr::select(pc)

# [1] "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"  "EC1V 4PY" "DE21 7BF" "E14 9SR"  "GL16 7YA" "WC1N 3XX"
# [10] "CB2 1AB" 
```

## Correct 1996-2010

The above exploratory analysis indicated that there is a huge increase in 2004-2006
in the above postcodes. I turn them to NAs and then use a regression to impute these
gaps.

```{r}
df9610 <- df9610 %>% 
  group_by(pc, year) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2003 & year <2007, NA, n))

df9610 <- pdata.frame(df9610)
predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
                                 model = "within", data = df9610), newdata = df9610), 2)

df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
```

## Load remaining data

```{r}
df9610 <- as_tibble(df9610) %>% 
  mutate(year = as.double(as.character(year))) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>%
  group_by(year, ladcd) %>%
  summarise(n = n()) %>%            
  ungroup()

@@

df1112 <- read_csv(data.path2011_2012) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  summarise(n = n()) %>%             # for n == 1  
  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
  ungroup()

df <- bind_rows(df9610, df1112) %>% 
  filter(!is.na(ladcd)) %>% 
  complete(ladcd, year) %>% 
  replace(is.na(.), 0)

@@

# [reproducible results]
# Random Forest 
# 
# 5984 samples
#    6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE        Rsquared   MAE       
#   2     variance    0.02336229  0.8713835  0.01536402
#   2     extratrees  0.02351745  0.8698132  0.01553849
#   4     variance    0.02405073  0.8667793  0.01599751
#   4     extratrees  0.02369748  0.8716644  0.01577096
#   6     variance    0.02504045  0.8591307  0.01680305
#   6     extratrees  0.02386677  0.8708794  0.01596582
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule =
#  variance and min.node.size = 5.

@@

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

@@

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

@@

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth_corrected.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

@@

stopCluster(cl)
end.time <- Sys.time() 
time.taken <- round(end.time - start.time,2)
time.taken # ca. 3 minutes 
```

@@

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth_corrected.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

@@
